R Notebook

### 2023-11-27 R_code/paper_analysis.Rmd
fix

```{r}
library(scales)
library(tidyverse)
library(caret)
library(ggpubr)
library(pROC)
library(predtools)
library(MASS)
library(caret)
```
```{r}
validation_data_prefix = "/Users/noa/Workspace/bootstrap_results/validation_data"
fasttree_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_fasttree.tsv", sep = ""))
raxml_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_raxml.tsv", sep = ""))
iqtree_validation_data = read_tsv(paste(validation_data_prefix,"/simulations_df_iqtree.tsv", sep = ""))
```

@@

fasttree_all_models_performance = read_tsv(paste(prefix,"ML_results/fasttree/all_models_performance.tsv", sep = ""))
fasttree_final_models_performance = read_tsv(paste(prefix,"ML_results/fasttree/final_model_performance.tsv", sep = ""))
fasttree_group_performance = read_tsv(paste(prefix,"ML_results/fasttree/groups_performance.tsv", sep = ""))
fasttree_vi =  read_tsv(paste(prefix,"ML_results/fasttree/final_model/full_model/model_standard_vi.tsv", sep = ""))
fasttree_enriched_test = read_tsv(paste(prefix,"ML_results/fasttree/test_with_predictions.tsv", sep = ""))
fasttree_validation = read_tsv(paste(prefix,"validation_data/simulations_df_fasttree.tsv", sep = ""))
#####
iqtree_all_models_performance = read_tsv(paste(prefix,"ML_results/iqtree/all_models_performance.tsv", sep = ""))
iqtree_final_models_performance = read_tsv(paste(prefix,"ML_results/iqtree/final_model_performance.tsv", sep = ""))
iqtree_group_performance = read_tsv(paste(prefix,"ML_results/iqtree/groups_performance.tsv", sep = ""))
iqtree_vi =  read_tsv(paste(prefix,"ML_results/iqtree/final_model/full_model/model_standard_vi.tsv", sep = ""))
iqtree_enriched_test = read_tsv(paste(prefix,"ML_results/iqtree/test_with_predictions.tsv", sep = ""))
iqtree_validation = read_tsv(paste(prefix,"validation_data/simulations_df_iqtree.tsv", sep = ""))
#####
raxml_all_models_performance = read_tsv(paste(prefix,"ML_results/raxml/all_models_performance.tsv", sep = ""))
raxml_final_models_performance = read_tsv(paste(prefix,"ML_results/raxml/final_model_performance.tsv", sep = ""))
raxml_group_performance = read_tsv(paste(prefix,"ML_results/raxml/groups_performance.tsv", sep = ""))
raxml_vi =  read_tsv(paste(prefix,"ML_results/raxml/final_model/full_model/model_standard_vi.tsv", sep = ""))
raxml_enriched_test = read_tsv(paste(prefix,"ML_results/raxml//test_with_predictions.tsv", sep = ""))
raxml_validation = read_tsv(paste(prefix,"validation_data/simulations_df_raxml.tsv", sep = ""))

Choose specific program
```{r}
program = 'fasttree'
if (program=='fasttree') {
    all_models_performance = fasttree_all_models_performance
    final_models_performance = fasttree_final_models_performance %>% filter(analysis_type!='feature_standard_fasttree_boot_support')
    group_performance = fasttree_group_performance
    vi = fasttree_vi
    enriched_test = fasttree_enriched_test
    #validation_data = fasttree_validation_data

} else if (program=='iqtree') {
    all_models_performance = iqtree_all_models_performance
    final_models_performance = iqtree_final_models_performance
    group_performance = iqtree_group_performance
    vi = iqtree_vi
    enriched_test = iqtree_enriched_test
    #validation_data = iqtree_validation_data
} else if (program=='raxml') {
    all_models_performance = raxml_all_models_performance
    final_models_performance = raxml_final_models_performance
    group_performance = raxml_group_performance
    vi = raxml_vi
    enriched_test = raxml_enriched_test
    #validation_data = raxml_validation
}
```

IQTREE
```{r}

summm_iqtree<- iqtree_enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_bayes = mean(feature_aBayes_iqtree_support), mean_ultrafast = mean(bootstrap_support), mean_bayes_cal = mean(predictions_feature_aBayes_iqtree_support_only_boot),
                                                mean_alrt = mean(feature_aLRT_iqtree_support), mean_alrt_cal = mean(predictions_feature_aLRT_iqtree_support_only_boot))

bayes<-summm_iqtree%>% ggplot(aes(y=mean_true,x=mean_bayes))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("aBayes test")+theme(text = element_text(size = 11))+ coord_flip()

alrt<-summm_iqtree %>% ggplot(aes(y=mean_true,x=mean_alrt))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("aLRT test ")+theme(text = element_text(size = 11))+ coord_flip()

ultrafast<-summm_iqtree %>% ggplot(aes(y=mean_true,x=mean_ultrafast/100))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("Ultrafast\nbootstrap")+theme(text = element_text(size = 11))+ coord_flip()

pred_iqtree<-summm_iqtree  %>% ggplot(aes(y=mean_true,x=mean_pred))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("IQTREE\nmodel")+theme(text = element_text(size = 11))+ coord_flip()

iqtree_plot<-ggarrange(pred_iqtree,ultrafast,bayes,alrt, labels = c("A","B","C"),align = "h", nrow = 2, ncol = 2 )
iqtree_plot
```
FASTTREE

```{r}
summm_fasttree<- fasttree_enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_sh = mean(bootstrap_support))

pred_fasttree<-summm_fasttree  %>% ggplot(aes(y=mean_true,x=mean_pred))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("Fasttree\nmodel")+theme(text = element_text(size = 11))+ coord_flip()

sh<-summm_fasttree  %>% ggplot(aes(y=mean_true,x=mean_sh))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("SH test")+theme(text = element_text(size = 11))+ coord_flip()

fasttree_plot<-ggarrange(pred_fasttree,sh, labels = c("A","B"),align = "h", nrow = 1, ncol = 2 )
fasttree_plot
```

RAXML
```{r}
summm_raxml<- raxml_enriched_test %>% group_by(tree_id) %>% summarise(mean_true = mean(true_binary_support), mean_pred = mean(prob_predictions), mean_sh = mean(bootstrap_support))

pred_raxml<-summm_raxml  %>% ggplot(aes(y=mean_true,x=mean_pred))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0, color = "green")+ylab("True fraction")+xlab("RAxML-NG\nmodel")+theme(text = element_text(size = 11))+ coord_flip()

rapid<-summm_fasttree  %>% ggplot(aes(y=mean_true,x=mean_sh))+geom_point(alpha = 1/3)+geom_abline(slope=1, intercept = 0,color = "green")+ylab("True fraction")+xlab("Bootstrap\nsupport")+theme(text = element_text(size = 11))+ coord_flip()

raxml_plot<-ggarrange(pred_raxml,rapid, labels = c("A","B"),align = "h", nrow = 1, ncol = 2 )
raxml_plot
```

```{r}
ggarrange(pred_iqtree,alrt, bayes, ultrafast,pred_raxml,rapid,"","",pred_fasttree,sh,align = "h", nrow = 3, ncol = 4 )

@@

paper_performance_iqtree<-iqtree_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,mcc_0.5,logloss, AUC,average_precision, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features", 
                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw ultrafast bootstrap support",
                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated ultrafast bootstrap support",
                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ ultrafast bootstrap support",
                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description,logloss, mcc_0.5,FPR,FNR, AUC,average_precision)  %>% arrange(dataset)

>>>>

```{r}
library(scales)
library(tidyverse)
library(caret)
library(ggpubr)
library(pROC)
library(predtools)
library(MASS)
library(caret)
```

@@

fasttree_all_models_performance = read_tsv(paste(prefix,"ML_results/fasttree/all_models_performance.tsv", sep = ""))
fasttree_final_models_performance = read_tsv(paste(prefix,"ML_results/fasttree/final_model_performance.tsv", sep = ""))
fasttree_group_performance = read_tsv(paste(prefix,"ML_results/fasttree/groups_performance.tsv", sep = ""))
fasttree_vi =  read_tsv(paste(prefix,"ML_results/fasttree/final_model/full_model/model_standard_vi.tsv", sep = ""))
fasttree_enriched_test = read_tsv(paste(prefix,"ML_results/fasttree/final_model/test.tsv", sep = ""))
fasttree_validation = read_tsv(paste(prefix,"validation_data/simulations_df_fasttree.tsv", sep = ""))
fasttree_enriched_val_downgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_downgrade.tsv", sep = ""))
fasttree_enriched_val_standard = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_upgrade.tsv", sep = ""))
fasttree_enriched_val_upgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_standard.tsv", sep = ""))
#####
iqtree_all_models_performance = read_tsv(paste(prefix,"ML_results/iqtree/all_models_performance.tsv", sep = ""))
iqtree_final_models_performance = read_tsv(paste(prefix,"ML_results/iqtree/final_model_performance.tsv", sep = ""))
iqtree_group_performance = read_tsv(paste(prefix,"ML_results/iqtree/groups_performance.tsv", sep = ""))
iqtree_vi =  read_tsv(paste(prefix,"ML_results/iqtree/final_model/full_model/model_standard_vi.tsv", sep = ""))
iqtree_enriched_test = read_tsv(paste(prefix,"ML_results/iqtree/final_model/test.tsv", sep = ""))
iqtree_enriched_val_downgrade = read_tsv(paste(prefix,"ML_results/iqtree/final_model/val_downgrade.tsv", sep = ""))
iqtree_enriched_val_standard = read_tsv(paste(prefix,"ML_results/iqtree/final_model/val_upgrade.tsv", sep = ""))
iqtree_enriched_val_upgrade = read_tsv(paste(prefix,"ML_results/iqtree/final_model/val_standard.tsv", sep = ""))
#####
raxml_all_models_performance = read_tsv(paste(prefix,"ML_results/raxml/all_models_performance.tsv", sep = ""))
raxml_final_models_performance = read_tsv(paste(prefix,"ML_results/raxml/final_model_performance.tsv", sep = ""))
raxml_group_performance = read_tsv(paste(prefix,"ML_results/raxml/groups_performance.tsv", sep = ""))
raxml_vi =  read_tsv(paste(prefix,"ML_results/raxml/final_model/full_model/model_standard_vi.tsv", sep = ""))
raxml_enriched_test = read_tsv(paste(prefix,"ML_results/raxml/final_model/test.tsv", sep = ""))
raxml_enriched_val_downgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_downgrade.tsv", sep = ""))
raxml_enriched_val_standard = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_upgrade.tsv", sep = ""))
raxml_enriched_val_upgrade = read_tsv(paste(prefix,"ML_results/fasttree/final_model/val_standard.tsv", sep = ""))

@@

paper_performance_iqtree<-iqtree_final_models_performance  %>% filter (metric_type=='all_data') %>% dplyr::select (dataset,analysis_type,name,brier_loss,logloss, AUC,average_precision,mcc_0.5, tp_0.5, tn_0.5, fp_0.5, fn_0.5)  %>% arrange(dataset) %>% mutate (description = case_when(analysis_type == "full_standard" & name =="model_standard" ~ "all features", 
                                analysis_type == "fast_standard" & name =="model_standard" ~ "All features except NNI features",
                                analysis_type == "bootstrap_support" & name =="raw_only_boot" ~ "Raw ultrafast bootstrap support",
                                analysis_type == "bootstrap_support" & name =="only_boot" ~ "Calibrated ultrafast bootstrap support",
                                analysis_type == "bootstrap_support" & name =="inc_boot" ~ "All features+ ultrafast bootstrap support",
                                    analysis_type == "feature_aLRT_iqtree_support" & name =="raw_only_boot" ~ "Raw aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="only_boot" ~ "Calibrated aLRT support",
                                analysis_type == "feature_aLRT_iqtree_support" & name =="inc_boot" ~ "All features+ aLRT support",
                                                                    analysis_type == "feature_aBayes_iqtree_support" & name =="raw_only_boot" ~ "Raw aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="only_boot" ~ "Calibrated aBayes support",
                                analysis_type == "feature_aBayes_iqtree_support" & name =="inc_boot" ~ "All features+ aBayes support",
                                )) %>% mutate(TPR =`tp_0.5`/(`tp_0.5`+`fn_0.5`),TNR =`tn_0.5`/(`tn_0.5`+`fp_0.5`),FPR =`fp_0.5`/(`tn_0.5`+`fp_0.5`),FNR =`fn_0.5`/(`tp_0.5`+`fn_0.5`)) %>% dplyr::select (dataset, description, brier_loss,logloss,AUC,average_precision, mcc_0.5,FPR,FNR, )  %>% arrange(dataset)
